---
title: "R Notebook"
output: html_notebook
---

# Critical Thinking 1
# 
#### Course Code: DS-610
#### Course Name: Advanced Applied Statistics
#### Module: 03
#
#### Student ID: G200007615
#### Student Name: Abdulaziz Alqumayzi
#### CRN: 15057

**Goal: Predict whether a song will reach a spot in the Top 10 of the Billboard Hot 100 Chart.**

Description of the variables:

- year = the year the song was released
- songtitle = the title of the song
- artistname = the name of the artist of the song
- songID and artistID = identifying variables for the song and artist
- timesignature and timesignature_confidence = a variable estimating the time signature of thesong, and the confidence in the estimate
- loudness = a continuous variable indicating the average amplitude of the audio in decibels
- tempo and tempo_confidence = a variable indicating the estimated beats per minute of thesong, and the confidence in the estimate
- key and key_confidence = a variable with twelve levels indicating the estimated key of the song(C, C#, . . ., B), and the confidence in the estimate
- energy = a variable that represents the overall acoustic energy of the song, using a mix offeatures such as loudness
- pitch = a continuous variable that indicates the pitch of the song
- timbre_0_min, timbre_0_max, timbre_1_min, timbre_1_max, . . . , timbre_11_min, andtimbre_11_max = variables that indicate the minimum/maximum values over all segments foreach of the twelve values in the timbre vector (resulting in 24 continuous variables)
- Top10 = a binary variable indicating whether or not the song made it to the Top 10 of theBillboard Hot 100 Chart (1 if it was in the top 10, and 0 if it was not)


#### Preprocessing 

```{r}
# import libraries
library(mlogit)
library(dfidx)
library(stats)

# read data 
songs <- read.csv("~/Master DS/Level-3/DS-610 Advanced Statistics/Critical Thinking/CT1/songs.csv")
View(songs)

# check duplicates 
length(unique(songs$songID)) == nrow(songs)

```

#### Tasks: 

Task 1:

a.Specify the total number of observations in the dataset

```{r}
# Specify the total number of observations
nrow(songs)

```
Number of observations are 7574.

b.Which artist has the most number of songs in the dataset?

```{r}
# the artist has the most number of songs in the dataset
sort(table(songs$artistname),decreasing=TRUE)[1:5]

```
The artist who has the most number of songs is Anal Cunt. since "Various artists" and "Various Artists" which indicate the same meaning are not an artist.

c.Split the data into training set (all songs before 2010) and test set (songs in 2010).

```{r}
# split data to training and testing sets 
train <- songs[songs$year < 2010,]
test <- songs[songs$year == 2010,]

```
d.How many samples are there in training set and test set respectively?

```{r}
# Specify the total number of observations in traning and testing sets
nrow(train)
nrow(test)

```

Task 2: 

Exclude the following variables from the model: "year", "songtitle", "artistname","songID", and "artistID"). 
Build a logistic regression model that predicts whether a song will be in the top 10.

```{r}
# excluding the following variables from the model: "year", "songtitle", "artistname","songID", and "artistID").
drops <- c("year","songtitle","artistname","songID","artistID")
train <- train[ , !(names(train) %in% drops)]
test <- test[ , !(names(test) %in% drops)]

# change Top10 variable type to factor 
songs$Top10 <- as.factor(songs$Top10)

# build a logistic regression model to predict whether a song will be in the top 10
glm_model <- glm(Top10 ~ timesignature + timesignature_confidence + loudness + tempo +
                  tempo_confidence + key + key_confidence + energy + pitch + timbre_0_min +
                  timbre_0_max + timbre_1_min + timbre_1_max + timbre_2_min + timbre_2_max + 
                  timbre_3_min + timbre_5_max + timbre_6_min + timbre_6_max + timbre_7_min +
                  timbre_7_max + timbre_8_min + timbre_8_max + timbre_9_min + timbre_9_max + 
                  timbre_10_min + timbre_10_max + timbre_11_min + timbre_11_max, data =  train,
                 family = binomial())


# predicting our model on the test dataset
predicted = predict(glm_model, newdata = test, type = 'response')
predicted

```

Task 3: Provide a table of your model summary

```{r}
# a table of the model summary
summary(glm_model)

```


Task 4: Investigating model summary:

a.Which parameters are not significant at 5% significance level. 

The parameters that don't have star, 2 starts or 3 stars sign on the right side of them.

The parameters: timesignature, tempo, timbre_1_max, timbre_2_max, timbre_3_min, timbre_5_max, timbre_8_min, timbre_8_max, timbre_9_min, timbre_9_max

b.Songs with heavier instrumentation tend to be louder (have higher values in the variable "loudness"). Do mainstream users prefer song with heavier or lighter instrumentation?

```{r}
# a summary of loudness variable
summary(songs$loudness)
```

The fact is whe the loudness vaiable near to zero means the sound is quite.
From the summary above, we can see the mean and median are between -7.6 and -8.8 values, the first quartile is -10.8 and the third quartile is -5.6. So most of the songs heard in this app are ranged between -5.6 and -8.8 decibels. That means users of this app are mostly prefer to hearing songs with lighter instruments. 

Task 5: Consider the 3 baseline models and provide accuracy score in test set for each model:

a.Predict each song to be in top 10

```{r}
# matrix model to predict each song to be in top 10 using 0.35 threshold
matrix_in_top10 <- table(Actual_value= test$Top10, Predicted_value= predicted >=0.35)
matrix_in_top10

```

The number of songs to be in the top 10 from the test dataset are 20 songs.

b.Predict each song to be not in top 10

```{r}
# simple model to predict each song not to be in top 10 using 0 threshold
not_in_top10 <- table(Actual_value= test$Top10, Predicted_value= predicted >=0)
not_in_top10

```

The number of songs to be not in the top 10 from the test dataset are 314 songs.

c.Randomly predict a song to be in top 10 with probability 0.2

```{r}
# matrix to randomly predict a song to be in top 10 with probability 0.2
table(Actual_value= test$Top10, Predicted_value= predicted >=0.2)

```

The number of songs to be in the top 10 using probability of 0.2 from the test dataset are 36 songs

d.Randomly predict a song to be in top 10 with probability 0.5

```{r}
# matrix to randomly predict a song to be in top 10 with probability 0.5
table(Actual_value= test$Top10, Predicted_value= predicted >=0.5)


```

The number of songs to be in the top 10 using probability of 0.5 from the test dataset are 10 songs

Task 6: Provide your model performance on the test set using the following 3 metrics (Use a different threshold ):

```{r}
# the model that we built 
matrix_in_top10
```

a. Accuracy score

```{r}
# accuracy score
((matrix_in_top10[[1,1]] + matrix_in_top10[[2,2]])/sum(matrix_in_top10))*100

```

Accuracy score is 86.32%

b. True Positive Rate

```{r}
# True Positive Rate
(matrix_in_top10[[2,2]]/(matrix_in_top10[[2,2]]+matrix_in_top10[[1,2]]))*100

```

The ratio of True Positive is 62.5%

c. False positive Rate

```{r}
# False positive Rate
(1-(matrix_in_top10[[2,2]]/(matrix_in_top10[[2,2]]+matrix_in_top10[[1,2]])))*100

```

The ratio of False Positive is 37.5%

Task 7: Provide your model performance on the test set using the following 3 metrics (Use threshold = 0.45):

```{r}
#matrix model to predict each song to be in top 10 using 0.45 threshold
matrix <- table(Actual_value= test$Top10, Predicted_value= predicted >=0.45)
matrix
```

a. Accuracy score
```{r}
# accuracy score
((matrix[[1,1]] + matrix[[2,2]])/sum(matrix))*100

```

Accuracy score is 87.13%

b. True Positive Rate

```{r}
# True Positive Rate
(matrix[[2,2]]/(matrix[[2,2]]+matrix[[1,2]]))*100

```

The ratio of True Positive is 92.3%

c. False positive Rate

```{r}
# False positive Rate
(1-(matrix[[2,2]]/(matrix[[2,2]]+matrix[[1,2]])))*100

```

The ratio of False Positive is 7.69%

**References:**

Ferrari, P. (2018, August 9). Logistic regression in R | logistic regression in R example | data science algorithms | simplilearn. YouTube. Retrieved October 2, 2021, from https://www.youtube.com/watch?v=XycruVLySDg. 

Chang, J. (2016). Predict the Top 10 of the Billboard. rstudio-pubs-static.s3.amazonaws.com. Retrieved October 2, 2021, from https://rstudio-pubs-static.s3.amazonaws.com/178029_dc0439885a694643a400fca032e7edb9.html. 

Field, A. P., Miles, J., &amp; Field Zoë. (2017). Discovering statistics using R. W. Ross MacDonald School Resource Services Library. 